Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features

### 2016-12-17 paper/structureScene.tex
Restructure expt #1 section
Change order of log and MFCCs vs scattering sections. Include 2016
results in the MFCCs vs scattering section. Remove GMM results for
MFCCs.

@@

\subsubsection*{Early \vs late integration}
Late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
This suggests that there is relevant variability in scales beyond the window size $T = 372~\mathrm{ms}$ that is not captured by averaging the scattering transform in time (early integration). One potential solution to this is to increase the scale of the second-order modulation wavelets from $J_2 = 12$ octaves to some larger value. This does not result in any performance gain, however (results not shown). What does help is the late integration, since it allows the SVM to successfully classify the few short extracts that characterize the scene which are then used to classify the whole scene through the majority vote. Since the scattering transform, even with larger values of $J_2$, is not able to separate out these short, discriminative sound events, delegating this task to the late-integrated results in a more powerful classification system.

\begin{table}
\begin{center}
\caption{Class-wise accuracies on the DCASE 2016 dataset for the baseline algorithm (left) and proposed algorithm (right).\label{table:dcase2016}}
\begin{tabular}{lccc}
Features & MFCCs & MFCCs & log-scattering \\
Classifier  & GMMs & linear SVMs & linear SVMs \\
Variant of \emph{late} integration & product & majority vote & majority vote \\
\midrule
beach & $74.6 \pm 18.9$ & $58.3 \pm 28.2$ & $83.5 \pm \phantom{0}7.1$ \\
bus & $58.2 \pm 18.3$ & $58.2 \pm 25.4$ & $88.8 \pm 11.4$ \\
cafe/restaurant & $85.1 \pm 10.8$ & $77.4 \pm 14.7$ & $64.5 \pm \phantom{0}8.3$ \\
car & $69.1 \pm 20.8$ & $92.3 \pm \phantom{0}7.7$ & $94.9 \pm \phantom{0}6.1$ \\
city\_center & $89.6 \pm \phantom{0}9.2$ & $96.4 \pm \phantom{0}3.8$ & $91.7 \pm \phantom{0}9.4$ \\
forest\_path & $72.4 \pm 23.8$ & $67.7 \pm 30.8$ & $93.8 \pm \phantom{0}7.8$ \\
grocery\_store & $74.1 \pm 13.7$ & $44.7 \pm 24.8$ & $90.9 \pm \phantom{0}6.9$ \\
home & $78.1 \pm 17.7$ & $44.8 \pm 15.3$ & $56.2 \pm 23.9$ \\
library & $65.1 \pm 23.3$ & $54.6 \pm 20.3$ & $82.0 \pm 13.0$ \\
metro\_station & $85.2 \pm 15.6$ & $94.7 \pm \phantom{0}6.8$ & $96.0 \pm \phantom{0}2.3$ \\
office & $90.8 \pm 16.0$ & $95.7 \pm \phantom{0}7.5$ & $87.5 \pm 21.7$ \\
park & $25.6 \pm 11.0$ & $52.4 \pm 10.4$ & $75.4 \pm \phantom{0}7.4$ \\
residential\_area & $75.1 \pm 19.4$ & $53.2 \pm 17.0$ & $44.2 \pm 15.0$ \\
train & $34.1 \pm \phantom{0}7.2$ & $44.5 \pm 23.1$ & $58.3 \pm 10.0$ \\
tram & $85.4 \pm 13.9$ & $52.1 \pm 11.9$ & $83.7 \pm 12.2$ \\
\bottomrule
Average & $70.8 \pm \phantom{0}2.6$ & $65.8 \pm \phantom{0}8.0$ & $79.4 \pm \phantom{0}3.0$ \\
\end{tabular}
\end{center}

\subsubsection*{Participation to the DCASE 2016 challenge}

In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
This edition relies on the TUT 2016 acoustic scenes dataset.

Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$, we therefore use a linear kernel instead.
Our experiments on the smaller DCASE 2013 dataset, as reported in Table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.

Our results on the public part of the DCASE 2016 dataset are reported in Table \ref{table:dcase2016}, compared with a baseline using MFCCs and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model. This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods. For comparison, we also provide the results for the same MFCC features coupled with a linear SVM classifier. Results here are slightly worse since the SVM can only discriminate between classes that are linearly separable, a limitation not present for the GMMs.

With the scattering transforms, however, the average accuracy across classes is brought from $70.8\%$ to $79.4\%$.\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.} It is interesting to note, however, that two classes, ``home'' and ``residential\_area'' both drop significantly in accuracy from the baseline GMM model to the scattering SVM classifier. This can be partially explained by the substitution of the GMM with an SVM, which shows a similar drop for the MFCCs. However, another reason could be that these classes are well-characterized by the average spectral envelope of the signal and the second-order modulation information captured by the scattering coefficients is therefore superfluous.
Furthermore, our cross-validation analysis confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.

\ml{this section takes one column and is I believe a good candidate for being shrunk or even merged with the previous subsection. By statement like, similar findings ... or those results are confirmed using the 2016 dataset ...}

>>>>

\begin{table}
\begin{center}
\caption{Class-wise accuracies on the development part of the DCASE 2016 dataset.\label{table:dcase2016}}
\begin{tabular}{lcc}
Features & MFCCs & log-scattering \\
\midrule
beach & $58.3 \pm 28.2$ & $83.5 \pm \phantom{0}7.1$ \\
bus & $58.2 \pm 25.4$ & $88.8 \pm 11.4$ \\
cafe/restaurant & $77.4 \pm 14.7$ & $64.5 \pm \phantom{0}8.3$ \\
car & $92.3 \pm \phantom{0}7.7$ & $94.9 \pm \phantom{0}6.1$ \\
city\_center & $96.4 \pm \phantom{0}3.8$ & $91.7 \pm \phantom{0}9.4$ \\
forest\_path & $67.7 \pm 30.8$ & $93.8 \pm \phantom{0}7.8$ \\
grocery\_store & $44.7 \pm 24.8$ & $90.9 \pm \phantom{0}6.9$ \\
home & $44.8 \pm 15.3$ & $56.2 \pm 23.9$ \\
library & $54.6 \pm 20.3$ & $82.0 \pm 13.0$ \\
metro\_station & $94.7 \pm \phantom{0}6.8$ & $96.0 \pm \phantom{0}2.3$ \\
office & $95.7 \pm \phantom{0}7.5$ & $87.5 \pm 21.7$ \\
park & $52.4 \pm 10.4$ & $75.4 \pm \phantom{0}7.4$ \\
residential\_area & $53.2 \pm 17.0$ & $44.2 \pm 15.0$ \\
train & $44.5 \pm 23.1$ & $58.3 \pm 10.0$ \\
tram & $52.1 \pm 11.9$ & $83.7 \pm 12.2$ \\
\bottomrule
Average & $65.8 \pm \phantom{0}8.0$ & $79.4 \pm \phantom{0}3.0$ \\
\end{tabular}
\end{center}

@@

In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge, which was evaluated on a much larger dataset compared to DCASE 2013. For this reason, training support vector machines with the Gaussian RBF kernel proved computationally too demanding and we instead used a linear kernel. Results on DCASE 2013 (see Table \ref{table:dcase2013}) suggests that this provides comparable, albeit slightly worse results.

We compare our results with baseline classifier using linear SVMs and MFCC features, which achieves an average accuracy of $65.5\%$. With the scattering transform, however, the average accuracy increases to $79.4\%$. This accuracy increases slightly to $80.8\%$ when calculated on the held-out evaluation dataset.\footnote{See \url{http://www.cs.tut.fi/sgn/arg/dcase2016/task-results-acoustic-scene-classification}} The small change in accuracy suggests that there is little overfitting in the model, an indication of good generalization power.

\subsubsection*{Early \vs late integration}
On the DCASE 2013 dataset, late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
This suggests that there is relevant variability in scales beyond the window size $T = 372~\mathrm{ms}$ that is not captured by averaging the scattering transform in time (early integration). One potential solution to this is to increase the scale of the second-order modulation wavelets from $J_2 = 12$ octaves to some larger value. This does not result in any performance gain, however (results not shown). What does help is the late integration, since it allows the SVM to successfully classify the few short extracts that characterize the scene which are then used to classify the whole scene through the majority vote. Since the scattering transform, even with larger values of $J_2$, is not able to separate out these short, discriminative sound events, delegating this task to the late-integrated results in a more powerful classification system.
